---
title: "AustinFamMidline_SurveyCleaning"
author: "Jasleen"
date: "2025-01-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Required Libraries

```{r load-libraries}
library(tidyverse)
library(magrittr)
```

## Define Variable Naming Conventions and Helper Functions

```{r define-helpers}
# Variable mapping for standardization
variable_mapping <- list(
  # Employment variables - prefix: emp_
  "work" = "emp_status",
  "work_hours_per_week" = "emp_hours",
  "work_hours_change" = "emp_hours_change",
  "work_increase" = "emp_increase",
  "work_decrease" = "emp_decrease",
  "work_no_reason" = "emp_unemployed_reason",
  "work_no_reason_care" = "emp_unemployed_care",
  
  # Housing variables - prefix: house_
  "housing_situation" = "house_status",
  "housing_burden" = "house_burden",
  "housing_behind" = "house_payments_behind",
  "moved_last_6mo" = "house_moved",
  "moved_last_6mo_num" = "house_moved_count",
  "move_reason" = "house_move_reason",
  
  # Wellbeing variables - keep wb_ prefix
  "wb_anxious" = "wb_anxious",
  "wb_worry" = "wb_worry",
  "wb_interest" = "wb_interest",
  "wb_down" = "wb_down",
  "wb_food_enough" = "wb_food_security",
  "wb_health_perception" = "wb_health_status",
  
  # Childcare variables - prefix: cc_
  "care_under_18" = "cc_has_children",
  "children_range" = "cc_age_range",
  "under_5_enroll" = "cc_under5_care",
  "above_5_enroll" = "cc_5to12_care",
  "above_12_enroll" = "cc_above12_care",
  "childcare_situation" = "cc_arrangement",
  "childcare_costs" = "cc_cost",
  
  # Financial & Goals - prefixes: fin_ and goal_
  "expenses_difficulty" = "fin_expense_difficulty",
  "spend_needs_meet" = "fin_needs_met",
  "fin_emergency" = "fin_emergency_fund",
  "goals_select" = "goal_priority",
  "goals_house" = "goal_housing",
  "goals_job" = "goal_employment",
  "goals_ed" = "goal_education"
)
# Helper function to handle "please specify" responses
handle_specify_responses <- function(df) {
  df %>%
    mutate(
      # Create new variable names for specify responses
      is_specify = str_detect(tolower(question_text), "(please specify)|(please describe)"),
      variable_name = if_else(is_specify,
                            paste0(variable_name, "_specify"),
                            variable_name)
    )
}

# Helper function to clean text responses
clean_text_responses <- function(x) {
  if(is.character(x)) {
    x %>%
      str_trim() %>%
      str_squish() %>%
      str_replace_all("[\\r\\n]", " ") %>%
      na_if("") %>%
      na_if("NA") %>%
      na_if("N/A")
  } else {
    x
  }
}

# Helper function to add timepoint suffix
add_timepoint <- function(df, timepoint = "m") {
  names(df) <- paste0(names(df), "_", timepoint)
  names(df) <- gsub(pattern = "responseid_[bm]$", 
                   replacement = "responseid", 
                   x = names(df))
  return(df)
}

# Helper function to standardize variable names
standardize_names <- function(df, mapping) {
  for(old_name in names(mapping)) {
    if(old_name %in% names(df)) {
      df <- df %>% rename(!!mapping[[old_name]] := old_name)
    }
  }
  return(df)
}
```

## Read and Initial Clean Data

```{r read-data}
# Read the CSV file
midline_data <- read_csv("austin_fam_midline.csv",
                        na = c("", "NA", "N/A", "#N/A", "NULL", "null"))

```

```{r,}
print("Initial data dimensions:")
print(dim(midline_data))

# Check the Finished column values
print("Values in Finished column:")
print(table(midline_data$Finished))

# Check initial data
print(paste("Initial rows:", nrow(midline_data)))

# Modified cleaning with proper Finished filter
midline_clean <- midline_data %>%
  # Remove the ImportId row and keep only TRUE/1
  filter(Finished %in% c("TRUE", "1", TRUE)) %>%
  {
    print(paste("Rows after Finished filter:", nrow(.)))
    .
  } %>%
  # Clean column names
  rename_all(~tolower(gsub(" ", "_", .))) %>%
  {
    print(paste("Rows after rename:", nrow(.)))
    .
  } %>%
  # Standardize variable names
  standardize_names(variable_mapping) %>%
  {
    print(paste("Rows after standardize:", nrow(.)))
    .
  } %>%
  # Add timepoint suffix
  add_timepoint("m") %>%
  {
    print(paste("Final rows:", nrow(.)))
    .
  }
```

```{r,}
# Modified cleaning with corrected filter
midline_clean <- midline_data %>%
  # Keep both TRUE and "Finished" rows, exclude only the ImportId row
  filter(!Finished == '{"ImportId":"finished"}') %>%
  {
    print(paste("Rows after Finished filter:", nrow(.)))
    .
  } %>%
  # Clean column names
  rename_all(~tolower(gsub(" ", "_", .))) %>%
  {
    print(paste("Rows after rename:", nrow(.)))
    .
  } %>%
  # Standardize variable names
  standardize_names(variable_mapping) %>%
  {
    print(paste("Rows after standardize:", nrow(.)))
    .
  } %>%
  # Add timepoint suffix
  add_timepoint("m") %>%
  {
    print(paste("Final rows:", nrow(.)))
    .
  }
```

## Clean Individual Sections

### Employment Section
```{r employment-cleaning}
# Employment Section Cleaning
emp_clean <- midline_clean %>%
  select(responseid, 
         starts_with("emp_"),
         contains("work")) %>%
  mutate(
    # Convert categorical to factors
    across(matches("emp_status|emp_increase|emp_decrease|emp_unemployed"), as.factor),
    # Handle numeric variables
    emp_hours_m = as.numeric(emp_hours_m),
    # Clean text responses
    across(ends_with(c("_text_m")), clean_text_responses),
    # Add logic checks
    emp_hours_valid = case_when(
      emp_status_m == "No" & !is.na(emp_hours_m) ~ FALSE,
      emp_hours_m > 168 ~ FALSE,
      TRUE ~ TRUE
    )
  )

# Let's examine the employment data quality
print("Employment Status Distribution:")
print(table(emp_clean$emp_status_m, useNA = "ifany"))

print("Hours Distribution:")
print(summary(emp_clean$emp_hours_m))

print("Number of Invalid Hours:")
print(sum(!emp_clean$emp_hours_valid, na.rm = TRUE))

# Check for problematic patterns
emp_problems <- emp_clean %>%
  filter(
    # People saying they don't work but reporting hours
    (emp_status_m == "No" & !is.na(emp_hours_m)) |
    # Unrealistic hours
    emp_hours_m > 168 |
    # Inconsistent unemployment reasons
    (emp_status_m == "Yes" & !is.na(emp_unemployed_reason_m))
  )

print("Number of problematic cases:")
print(nrow(emp_problems))
```

# Housing Section Cleaning
```{r housing-clean}
# Housing Section Cleaning - Modified to handle move counts correctly
housing_clean <- midline_clean %>%
  select(responseid, 
         starts_with("house_"),
         contains("move")) %>%
  mutate(
    # Clean housing status
    house_status_m = case_when(
      str_detect(house_status_m, "^We honor") ~ NA_character_,
      TRUE ~ house_status_m
    ),
    house_status_m = as.factor(house_status_m),
    
    # Convert other categoricals to factors
    across(c(house_burden_m, house_payments_behind_m, 
             house_moved_m), as.factor),
             
    # Convert move count from text to numeric
    house_moved_count_clean = case_when(
      house_moved_count_m == "1 time" ~ 1,
      house_moved_count_m == "2 times" ~ 2,
      house_moved_count_m == "3 or more times" ~ 3,
      str_detect(house_moved_count_m, "^How many times") ~ NA_real_,
      TRUE ~ NA_real_
    ),
    
    # Add validity check
    moves_valid = case_when(
      house_moved_m == "No" & !is.na(house_moved_count_clean) ~ FALSE,
      TRUE ~ TRUE
    )
  )

# Check cleaned data
print("Housing Status Distribution (cleaned):")
print(table(housing_clean$house_status_m, useNA = "ifany"))

print("Move Count Distribution (cleaned):")
print(table(housing_clean$house_moved_count_clean, useNA = "ifany"))

print("Number of Invalid Cases:")
print(sum(!housing_clean$moves_valid, na.rm = TRUE))
```

# wellbeing data cleaning
```{r wellbeing-cleaning}
wellbeing_clean <- midline_clean %>%
  select(responseid, 
         starts_with("wb_")) %>%
  mutate(
    # Clean anxiety and depression measures
    across(c(wb_anxious_m, wb_worry_m, wb_interest_m, wb_down_m), 
           ~case_when(
             str_detect(., "^Over the last") ~ NA_character_,
             TRUE ~ .
           )),
    
    # Clean food security
    wb_food_security_m = case_when(
      str_detect(wb_food_security_m, "^Getting enough food") ~ NA_character_,
      TRUE ~ wb_food_security_m
    ),
    
    # Clean health status
    wb_health_status_m = case_when(
      str_detect(wb_health_status_m, "^Would you say") ~ NA_character_,
      TRUE ~ wb_health_status_m
    ),
    
    # Convert all to factors
    across(starts_with("wb_"), as.factor),
    
    # Add validation flags
    wb_mental_health_complete = !if_any(c(wb_anxious_m, wb_worry_m, wb_interest_m, wb_down_m), is.na),
    wb_food_health_complete = !if_any(c(wb_food_security_m, wb_health_status_m), is.na)
  )

# Check cleaned distributions
print("MENTAL HEALTH MEASURES:")
for(col in c("wb_anxious_m", "wb_worry_m", "wb_interest_m", "wb_down_m")) {
  print(paste("\n", col, ":"))
  print(table(wellbeing_clean[[col]], useNA = "ifany"))
}

print("\nFOOD SECURITY AND HEALTH STATUS:")
print("\nFood Security:")
print(table(wellbeing_clean$wb_food_security_m, useNA = "ifany"))
print("\nHealth Status:")
print(table(wellbeing_clean$wb_health_status_m, useNA = "ifany"))

print("\nCompletion Rates:")
print(paste("Mental health measures complete:", sum(wellbeing_clean$wb_mental_health_complete)))
print(paste("Food/health measures complete:", sum(wellbeing_clean$wb_food_health_complete)))
```

# Childcare Section Cleaning
```{r childcare-cleaning}
childcare_clean <- midline_clean %>%
  select(responseid, 
         starts_with("cc_"),
         contains("child"),
         contains("care")) %>%
  mutate(
    # Clean has_children responses
    cc_has_children_m = case_when(
      str_detect(cc_has_children_m, "^Families take") ~ NA_character_,
      TRUE ~ cc_has_children_m
    ),
    
    # Clean age range but maintain original categories
    cc_age_range_m = case_when(
      str_detect(cc_age_range_m, "^For the kids") ~ NA_character_,
      TRUE ~ cc_age_range_m
    ),
    
    # Clean childcare arrangement
    cc_arrangement_m = case_when(
      str_detect(cc_arrangement_m, "^Over the past") ~ NA_character_,
      TRUE ~ cc_arrangement_m
    ),
    
    # Convert to factors
    cc_has_children_m = as.factor(cc_has_children_m),
    cc_age_range_m = as.factor(cc_age_range_m),
    cc_arrangement_m = as.factor(cc_arrangement_m)
  )

# Check distributions
print("Has Children Distribution (cleaned):")
print(table(childcare_clean$cc_has_children_m, useNA = "ifany"))

print("\nAge Range Distribution (original categories):")
print(table(childcare_clean$cc_age_range_m[childcare_clean$cc_has_children_m == "Yes"], 
            useNA = "ifany"))

print("\nChildcare Arrangement:")
print(table(childcare_clean$cc_arrangement_m[childcare_clean$cc_has_children_m == "Yes"], 
            useNA = "ifany"))

childcare_clean <- childcare_clean %>%
  mutate(
    # Add validity checks
    valid_response = case_when(
      is.na(cc_has_children_m) ~ FALSE,
      cc_has_children_m == "Yes" & is.na(cc_age_range_m) ~ FALSE,
      cc_has_children_m == "Yes" & is.na(cc_arrangement_m) ~ FALSE,
      cc_has_children_m == "No" & !is.na(cc_age_range_m) ~ FALSE,
      cc_has_children_m == "No" & !is.na(cc_arrangement_m) ~ FALSE,
      TRUE ~ TRUE
    )
  )

# Check validity
print("\nValidity Summary:")
print(paste("Number of invalid cases:", sum(!childcare_clean$valid_response)))
print(paste("Percent valid:", round(mean(childcare_clean$valid_response) * 100, 1), "%"))
```

# Financial and Impact Section Cleaning
```{r financial-income cleaning}
financial_clean <- midline_clean %>%
  select(responseid, 
         starts_with("fin_"),
         contains("expense"),
         contains("benefit"),
         contains("emergency"),
         contains("needs"),
         contains("spend")) %>%
  mutate(
    # Clean expense difficulty
    fin_expense_difficulty_m = case_when(
      str_detect(fin_expense_difficulty_m, "^In the last month") ~ NA_character_,
      TRUE ~ fin_expense_difficulty_m
    ),
    
    # Clean emergency fund response
    fin_emergency_fund_m = case_when(
      str_detect(fin_emergency_fund_m, "^In the past") ~ NA_character_,
      TRUE ~ fin_emergency_fund_m
    ),
    
    # Create binary indicators for emergency management strategies
    emergency_borrowed_family = ifelse(is.na(fin_emergency_select_m), FALSE, 
                                     str_detect(fin_emergency_select_m, "borrowed money from family")),
    emergency_reduced_spending = ifelse(is.na(fin_emergency_select_m), FALSE,
                                      str_detect(fin_emergency_select_m, "reduce spending")),
    emergency_loan_credit = ifelse(is.na(fin_emergency_select_m), FALSE,
                                 str_detect(fin_emergency_select_m, "loan or used a credit card")),
    emergency_used_savings = ifelse(is.na(fin_emergency_select_m), FALSE,
                                  str_detect(fin_emergency_select_m, "emergency fund")),
    emergency_used_uptogether = ifelse(is.na(fin_emergency_select_m), FALSE,
                                     str_detect(fin_emergency_select_m, "UpTogether investment")),
    
    # Create binary indicators for needs met sources
    needs_regular_income = ifelse(is.na(fin_needs_met_m), FALSE,
                                str_detect(fin_needs_met_m, "Regular income sources")),
    needs_credit = ifelse(is.na(fin_needs_met_m), FALSE,
                         str_detect(fin_needs_met_m, "Credit cards or loans")),
    needs_savings = ifelse(is.na(fin_needs_met_m), FALSE,
                          str_detect(fin_needs_met_m, "Money from savings")),
    needs_borrowing = ifelse(is.na(fin_needs_met_m), FALSE,
                            str_detect(fin_needs_met_m, "Borrowing from friends")),
    needs_snap = ifelse(is.na(fin_needs_met_m), FALSE,
                       str_detect(fin_needs_met_m, "SNAP")),
    needs_wic = ifelse(is.na(fin_needs_met_m), FALSE,
                      str_detect(fin_needs_met_m, "WIC")),
    needs_school_meals = ifelse(is.na(fin_needs_met_m), FALSE,
                               str_detect(fin_needs_met_m, "School meal")),
    needs_rental = ifelse(is.na(fin_needs_met_m), FALSE,
                         str_detect(fin_needs_met_m, "Rental Assistance")),
    needs_unemployment = ifelse(is.na(fin_needs_met_m), FALSE,
                              str_detect(fin_needs_met_m, "Unemployment insurance")),
    needs_other = ifelse(is.na(fin_needs_met_m), FALSE,
                        str_detect(fin_needs_met_m, "Other")),
    
    # Convert to factors
    fin_expense_difficulty_m = as.factor(fin_expense_difficulty_m),
    fin_emergency_fund_m = as.factor(fin_emergency_fund_m)
  )

# Check distributions with na.rm=TRUE
print("Emergency Management Strategies (% of those with emergencies):")
print(paste("Borrowed from family/friends:", 
            round(mean(financial_clean$emergency_borrowed_family[financial_clean$fin_emergency_fund_m == "Yes"], na.rm=TRUE) * 100, 1), "%"))
print(paste("Reduced spending:", 
            round(mean(financial_clean$emergency_reduced_spending[financial_clean$fin_emergency_fund_m == "Yes"], na.rm=TRUE) * 100, 1), "%"))
print(paste("Used loan/credit:", 
            round(mean(financial_clean$emergency_loan_credit[financial_clean$fin_emergency_fund_m == "Yes"], na.rm=TRUE) * 100, 1), "%"))
print(paste("Used emergency fund:", 
            round(mean(financial_clean$emergency_used_savings[financial_clean$fin_emergency_fund_m == "Yes"], na.rm=TRUE) * 100, 1), "%"))
print(paste("Used UpTogether:", 
            round(mean(financial_clean$emergency_used_uptogether[financial_clean$fin_emergency_fund_m == "Yes"], na.rm=TRUE) * 100, 1), "%"))

print("\nNeeds Met Sources (% of all respondents):")
print(paste("Regular income:", round(mean(financial_clean$needs_regular_income, na.rm=TRUE) * 100, 1), "%"))
print(paste("Credit/loans:", round(mean(financial_clean$needs_credit, na.rm=TRUE) * 100, 1), "%"))
print(paste("Savings:", round(mean(financial_clean$needs_savings, na.rm=TRUE) * 100, 1), "%"))
print(paste("Borrowing:", round(mean(financial_clean$needs_borrowing, na.rm=TRUE) * 100, 1), "%"))
print(paste("SNAP:", round(mean(financial_clean$needs_snap, na.rm=TRUE) * 100, 1), "%"))
print(paste("WIC:", round(mean(financial_clean$needs_wic, na.rm=TRUE) * 100, 1), "%"))
print(paste("School meals:", round(mean(financial_clean$needs_school_meals, na.rm=TRUE) * 100, 1), "%"))
print(paste("Rental assistance:", round(mean(financial_clean$needs_rental, na.rm=TRUE) * 100, 1), "%"))
print(paste("Unemployment:", round(mean(financial_clean$needs_unemployment, na.rm=TRUE) * 100, 1), "%"))
print(paste("Other:", round(mean(financial_clean$needs_other, na.rm=TRUE) * 100, 1), "%"))
```
# Cleaning Goals Section
```{r goals}
goals_clean <- midline_clean %>%
  select(responseid, 
         starts_with("goal_"),
         starts_with("goals_")) %>%
  mutate(
    # Clean main goals question
    goals_m = case_when(
      str_detect(goals_m, "^At this point") ~ NA_character_,
      TRUE ~ goals_m
    ),
    
    # Create binary indicators for each goal type
    goal_housing = str_detect(goal_priority_m, "housing situation"),
    goal_employment = str_detect(goal_priority_m, "employment status"),
    goal_education = str_detect(goal_priority_m, "education"),
    goal_debt = str_detect(goal_priority_m, "debt"),
    goal_savings = str_detect(goal_priority_m, "Saving for the future"),
    goal_transport = str_detect(goal_priority_m, "transportation"),
    goal_other = str_detect(goal_priority_m, "Other"),
    
    # Convert main goals to factor
    goals_m = as.factor(goals_m),
    
    # Add validity check
    goals_response_valid = !is.na(goals_m),
    has_priority_goals = if_else(goals_m == "Yes", 
                               !is.na(goal_priority_m), 
                               TRUE)
  )

# Check distributions
print("Main Goals Distribution:")
print(table(goals_clean$goals_m, useNA = "ifany"))

print("\nType of Goals Selected (among those with goals):")
goals_summary <- goals_clean %>%
  filter(goals_m == "Yes") %>%
  summarise(
    Housing = sum(goal_housing),
    Employment = sum(goal_employment),
    Education = sum(goal_education),
    Debt = sum(goal_debt),
    Savings = sum(goal_savings),
    Transportation = sum(goal_transport),
    Other = sum(goal_other)
  )
print(goals_summary)

print("\nValidity:")
print(paste("Valid responses:", sum(goals_clean$goals_response_valid)))
print(paste("Has priority goals (when expected):", sum(goals_clean$has_priority_goals, na.rm = TRUE)))
```
# Community and Social capital section cleaning
```{r Community-social capital}
community_clean <- midline_clean %>%
  select(responseid,
         starts_with("comm_"),
         starts_with("ripple_"),
         starts_with("impact_ripple")) %>%
  mutate(
    # Clean and convert to factors
    across(starts_with("comm_give"), ~case_when(
      . %in% c("Never", "Sometimes", "Often") ~ .,
      TRUE ~ NA_character_
    )),
    
    # Convert to numeric scores (0-2)
    across(starts_with("comm_give"), ~case_when(
      . == "Never" ~ 0L,
      . == "Sometimes" ~ 1L,
      . == "Often" ~ 2L,
      TRUE ~ NA_integer_
    ), .names = "{.col}_score"),
    
    # Calculate total support score
    support_score_total = rowSums(across(ends_with("_score")), na.rm = TRUE),
    
    # Create binary indicators (TRUE/FALSE)
    gives_emotional = ifelse(!is.na(comm_give_emo_supp_m), 
                           comm_give_emo_supp_m %in% c("Often", "Sometimes"), 
                           NA),
    
    gives_material = ifelse(!is.na(comm_give_money_m) & !is.na(comm_give_food_m),
                           (comm_give_money_m %in% c("Often", "Sometimes") | 
                            comm_give_food_m %in% c("Often", "Sometimes")),
                           NA),
    
    gives_practical = ifelse(!is.na(comm_give_transport_m) & !is.na(comm_give_errands_m),
                            (comm_give_transport_m %in% c("Often", "Sometimes") | 
                             comm_give_errands_m %in% c("Often", "Sometimes")),
                            NA),
    
    # Ripple effects (TRUE/FALSE)
    supports_family = ripple_fam_select_m == "Yes",
    supports_community = ripple_comm_select_m == "Yes",
    
    # Convert original variables to factors
    across(c(starts_with("comm_give"), 
             ends_with("select_m")), as.factor)
  )

# Check distributions
print("Support Score Summary:")
print(summary(community_clean$support_score_total))

print("\nSupport Types (% giving sometimes/often):")
print(paste("Emotional support:", 
            round(mean(community_clean$gives_emotional, na.rm=TRUE) * 100), "%"))
print(paste("Material support:", 
            round(mean(community_clean$gives_material, na.rm=TRUE) * 100), "%"))
print(paste("Practical support:", 
            round(mean(community_clean$gives_practical, na.rm=TRUE) * 100), "%"))

print("\nRipple Effects:")
print(table(community_clean$supports_family, useNA = "ifany"))
print(table(community_clean$supports_community, useNA = "ifany"))
```
# Cleaned dataset
```{r final-cleaned dataset}

# Output the final merged dataset
write.csv(midline_clean, "midline_survey_final_cleaned.csv", row.names = FALSE)

# Print out some information about the final dataset
cat("Midline Survey Clean Dataset\n")
cat("Total number of respondents:", nrow(midline_clean), "\n")
cat("Number of columns in dataset:", ncol(midline_clean), "\n")
```

