---
title: "midline_descriptive_statistics"
author: "Jasleen"
date: "2025-01-12"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: united
    highlight: tango
    code_folding: hide
    fig_caption: true
    df_print: paged
  pdf_document:
    toc: true
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,          # Don't show code
  message = FALSE,       # Don't show messages
  warning = FALSE,       # Don't show warnings
  fig.align = "center",  # Center figures
  out.width = "80%",     # Control figure width
  fig.width = 10,        # Set figure width
  fig.height = 6         # Set figure height
)
```

```{r libraries}
library(tidyverse)
library(knitr)
```

```{r datareading}
setwd("/Users/marinatolchinsky/Library/CloudStorage/GoogleDrive-marina@uptogether.org/My Drive/Analytics/Hiring, Onboarding & PD/Data for Survey Analyst (Contract)/Jasleen")
data <- read.csv("midline_survey_final_cleaned.csv", stringsAsFactors = FALSE)
```
## Employment {.tabset}
Analysis of employment status, work hours, and reasons for unemployment among fund members.
```{r employment}
## Employment Analysis
# Function to calculate rounded percentages
calculate_percentage <- function(count, total) {
  round(count / total * 100, 0)
}

# Set default theme for all plots with larger fonts
theme_larger_text <- theme_minimal() +
  theme(
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 14),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12),
    strip.text = element_text(size = 12, face = "bold")
  )

theme_set(theme_larger_text)

# 1. Employment Status Distribution
emp_status_df <- data.frame(
  Employment_Status = c("Employed", "Unemployed", "Preferred not to answer"),
  Count = c(
    sum(data$emp_status_m == "Yes", na.rm = TRUE),
    sum(data$emp_status_m == "No", na.rm = TRUE),
    sum(data$emp_status_m == "Prefer not to answer", na.rm = TRUE)
  )
)

total_responses <- sum(emp_status_df$Count)
emp_status_df$Percentage <- calculate_percentage(emp_status_df$Count, total_responses)

# Display the formatted table with sample size
cat(sprintf("\nEmployment Status Distribution (n=%d):\n", total_responses))
print(knitr::kable(emp_status_df,
      format = "simple",
      col.names = c("Employment Status", "Count", "Percentage (%)"),
      align = c('l', 'r', 'r')))

# Key findings text summary with larger text
cat("\nKey Employment Findings:\n")
cat(sprintf("- %d%% of fund members are employed (%d individuals)\n", 
    emp_status_df$Percentage[1], emp_status_df$Count[1]))
cat(sprintf("- %d%% of fund members are unemployed (%d individuals)\n", 
    emp_status_df$Percentage[2], emp_status_df$Count[2]))
cat(sprintf("- %d%% preferred not to answer (%d individuals)\n", 
    emp_status_df$Percentage[3], emp_status_df$Count[3]))

# Analyze unemployment reasons
unemployed_count <- sum(data$emp_status_m == "No", na.rm = TRUE)
unemployment_reasons_df <- data.frame(
  table(data$emp_unemployed_reason_m[data$emp_status_m == "No"])
)
names(unemployment_reasons_df) <- c("Reason", "Count")
unemployment_reasons_df$Percentage <- calculate_percentage(unemployment_reasons_df$Count, 
                                                        sum(unemployment_reasons_df$Count))

# Display unemployment reasons with sample size
cat(sprintf("\nReasons for Unemployment (n=%d):\n", unemployed_count))
print(knitr::kable(unemployment_reasons_df,
      format = "simple",
      col.names = c("Reason", "Count", "Percentage (%)"),
      align = c('l', 'r', 'r')))

# Work hours for employed individuals
employed_count <- sum(data$emp_status_m == "Yes", na.rm = TRUE)
work_hours_df <- data.frame(
  table(data$emp_hours_m[data$emp_status_m == "Yes"])
)
names(work_hours_df) <- c("Hours", "Count")
work_hours_df$Percentage <- calculate_percentage(work_hours_df$Count, 
                                              sum(work_hours_df$Count))

# Display work hours with sample size
cat(sprintf("\nWork Hours Distribution (n=%d employed individuals):\n", employed_count))
print(knitr::kable(work_hours_df,
      format = "simple",
      col.names = c("Hours Category", "Count", "Percentage (%)"),
      align = c('l', 'r', 'r')))

# Hours change analysis
hours_change_df <- data.frame(
  table(data$emp_hours_change_m[data$emp_status_m == "Yes"])
)
names(hours_change_df) <- c("Change", "Count")
hours_change_df$Percentage <- calculate_percentage(hours_change_df$Count, 
                                                sum(hours_change_df$Count))
table(data$emp_decrease_m)

# Display hours change with sample size
cat(sprintf("\nChanges in Work Hours (Past 6 Months) (n=%d employed individuals):\n", 
            employed_count))
print(knitr::kable(hours_change_df,
      format = "simple",
      col.names = c("Change in Hours", "Count", "Percentage (%)"),
      align = c('l', 'r', 'r')))

# Create a categorized summary of "other" reasons
reasons_categorized <- data.frame(
  Category = c(
    "Health Limitations",
    "Childcare/Family Responsibilities", 
    "Loss of Previous Job/Client",
    "Semi-Retired/Seeking Remote Work",
    "Other Personal Circumstances"
  ),
  Count = c(2, 3, 2, 1, 1)
)

# Calculate percentages with rounding
total_other_reasons <- sum(reasons_categorized$Count)
reasons_categorized$Percentage <- calculate_percentage(reasons_categorized$Count, 
                                                    total_other_reasons)

# Create simplified table with sample size
cat(sprintf("\nCategorized 'Other' Reasons for Unemployment (n=%d):\n", total_other_reasons))
print(knitr::kable(reasons_categorized,
      col.names = c("Category", "Count", "Percentage (%)"),
      align = c('l', 'r', 'r'),
      format = "simple"))

# Create visualization with sample size in title and larger font sizes
ggplot(reasons_categorized, 
       aes(x = reorder(Category, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "#46214A") +
  coord_flip() +
  labs(
    title = sprintf("Breakdown of 'Other' Reasons for Unemployment (n=%d)", total_other_reasons),
    x = "",
    y = "Number of Responses"
  ) +
  theme_larger_text +
  theme(
    axis.text.y = element_text(size = 12),  # Specific size for category labels
    axis.text.x = element_text(size = 12),  # Specific size for count labels
    plot.margin = margin(20, 20, 20, 20)    # Add more margin around the plot
  )

# Key findings text
cat("\nKey Findings from 'Other' Unemployment Reasons:\n")
cat("- Primary reasons include childcare responsibilities, health limitations, and loss of previous employment")
cat("- Some respondents are seeking specific work arrangements (e.g., remote work)")
cat("- Several responses indicate compounding challenges (e.g., childcare + job search difficulties)")
```
## Income and Affordability {.tabset}
Examination of household income, financial difficulties, and methods of meeting expenses.
```{r income-affordability}
# Create binary indicators for needs met sources
data <- data %>%
  mutate(
    needs_met_binary = case_when(
      !is.na(fin_needs_met_m) ~ 1,
      TRUE ~ 0
    )
  )

# 1. Household Income Distribution
total_responses_income <- sum(!is.na(data$fin_expense_difficulty_m))
income_dist <- data %>%
  filter(!is.na(fin_expense_difficulty_m)) %>%
  count(fin_expense_difficulty_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nHousehold Income Distribution (n=%d):\n", total_responses_income))
print(knitr::kable(income_dist,
      col.names = c("Difficulty Level", "Count", "Percentage (%)"),
      format = "simple"))

# 2. Methods of Meeting Spending Needs - Modified for multiple selections with WIC handling
total_responses_needs <- sum(!is.na(data$fin_needs_met_m))

# Function to properly split responses while preserving WIC
split_preserve_wic <- function(x) {
  # Replace the comma in WIC temporarily
  x <- gsub("Women, Infants, and Children", "Women_Infants_and_Children", x)
  # Split by comma and space
  parts <- strsplit(x, ",\\s*")[[1]]
  # Restore the original WIC text
  parts <- gsub("Women_Infants_and_Children", "Women, Infants, and Children", parts)
  return(parts)
}

# Process each non-NA response
spending_methods_list <- lapply(data$fin_needs_met_m[!is.na(data$fin_needs_met_m)], split_preserve_wic)
all_methods <- unlist(spending_methods_list)
spending_methods <- data.frame(
  Method = names(table(all_methods)),
  Count = as.numeric(table(all_methods))
) %>%
  mutate(
    percentage = round(Count/total_responses_needs * 100, 0)
  ) %>%
  arrange(desc(percentage))

cat(sprintf("\nMethods Used to Meet Spending Needs (n=%d respondents, multiple selections allowed):\n", 
            total_responses_needs))
print(knitr::kable(spending_methods,
      col.names = c("Method", "Count", "Percentage of Respondents (%)"),
      format = "simple"))

# 3. Benefits Reduction Analysis
total_responses_benefits <- sum(!is.na(data$benefits_reduce_m))
benefits_reduction <- data %>%
  filter(!is.na(benefits_reduce_m)) %>%
  count(benefits_reduce_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nExperience with Benefits Reduction (n=%d):\n", total_responses_benefits))
print(knitr::kable(benefits_reduction,
      col.names = c("Benefits Status", "Count", "Percentage (%)"),
      format = "simple"))

# 4. Financial Emergencies
total_responses_emergency <- sum(!is.na(data$fin_emergency_fund_m))
fin_emergency <- data %>%
  filter(!is.na(fin_emergency_fund_m)) %>%
  count(fin_emergency_fund_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nFinancial Emergency Experience (n=%d):\n", total_responses_emergency))
print(knitr::kable(fin_emergency,
      col.names = c("Experienced Emergency", "Count", "Percentage (%)"),
      format = "simple"))

# Key Summary Statistics with corrected metrics
summary_stats <- data.frame(
  Metric = c(
    "Households Reporting Difficulty with Expenses",
    "Households Using Benefits",
    "Households Experiencing Financial Emergency",
    "Households with Benefits Reduction"
  ),
  Percentage = c(
    round(mean(data$fin_expense_difficulty_m %in% 
              c("Very difficult", "Somewhat difficult"), na.rm = TRUE) * 100, 0),
    round(mean(data$needs_met_binary, na.rm = TRUE) * 100, 0),
    round(mean(data$fin_emergency_fund_m == "Yes", na.rm = TRUE) * 100, 0),
    round(mean(grepl("Yes", data$benefits_reduce_m), na.rm = TRUE) * 100, 0)
  )
)

summary_stats$Sample_Size <- c(
  sum(!is.na(data$fin_expense_difficulty_m)),
  sum(!is.na(data$fin_needs_met_m)),
  sum(!is.na(data$fin_emergency_fund_m)),
  sum(!is.na(data$benefits_reduce_m))
)

cat("\nKey Income and Affordability Metrics:\n")
print(knitr::kable(summary_stats,
      col.names = c("Metric", "Percentage (%)", "Sample Size (n)"),
      format = "simple"))
```
## Housing {.tabset}
Overview of housing status, cost burden, and mobility patterns.
```{r housing}
# 1. Housing Status Distribution with corrected categories
total_housing_responses <- sum(!is.na(data$house_status_m))

housing_status <- data %>%
  filter(!is.na(house_status_m)) %>%
  mutate(
    house_status_m = case_when(
      house_status_m == "No fixed residence (e.g., unsheltered, other places not made for housing)" ~ "Homeless",
      house_status_m == "Temporary housing (e.g., shelter, transitional housing)" ~ "Homeless",
      TRUE ~ house_status_m
    )
  ) %>%
  count(house_status_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  ) %>%
  arrange(desc(percentage))

cat(sprintf("\nHousing Status Distribution (n=%d):\n", total_housing_responses))
print(knitr::kable(housing_status,
      col.names = c("Housing Status", "Count", "Percentage (%)"),
      format = "simple"))

# 2. Housing Cost Burden
total_burden_responses <- sum(!is.na(data$house_burden_m))
cost_burden <- data %>%
  filter(!is.na(house_burden_m)) %>%
  count(house_burden_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nHousing Cost Burden (n=%d):\n", total_burden_responses))
print(knitr::kable(cost_burden,
      col.names = c("Cost Burden Level", "Count", "Percentage (%)"),
      format = "simple"))

# 3. Payment Status
total_payment_responses <- sum(!is.na(data$house_payments_behind_m))
payment_status <- data %>%
  filter(!is.na(house_payments_behind_m)) %>%
  count(house_payments_behind_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nHousing Payment Status (n=%d):\n", total_payment_responses))
print(knitr::kable(payment_status,
      col.names = c("Payment Status", "Count", "Percentage (%)"),
      format = "simple"))

# 4. Moving Patterns
total_moving_responses <- sum(!is.na(data$house_moved_m))
moving_status <- data %>%
  filter(!is.na(house_moved_m)) %>%
  count(house_moved_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nMoved in Last 6 Months (n=%d):\n", total_moving_responses))
print(knitr::kable(moving_status,
      col.names = c("Moving Status", "Count", "Percentage (%)"),
      format = "simple"))

# Key Summary Statistics with corrected calculations
summary_stats <- data.frame(
  Metric = c(
    "Renters",
    "Homeowners",
    "Experiencing Homelessness",
    "Severely Cost Burdened (>50% income on housing)",
    "Moved in Past 6 Months",
    "Behind on Payments"
  ),
  Percentage = c(
    round(mean(data$house_status_m == "Rented", na.rm = TRUE) * 100, 0),
    round(mean(stringr::str_detect(data$house_status_m, "Owned by you"), na.rm = TRUE) * 100, 0),
    round(mean(data$house_status_m %in% c(
      "No fixed residence (e.g., unsheltered, other places not made for housing)",
      "Temporary housing (e.g., shelter, transitional housing)"
    ), na.rm = TRUE) * 100, 0),
    round(mean(data$house_burden_m == "More than 50% of monthly income", na.rm = TRUE) * 100, 0),
    round(mean(data$house_moved_m == "Yes", na.rm = TRUE) * 100, 0),
    round(mean(data$house_payments_behind_m == "Yes", na.rm = TRUE) * 100, 0)
  )
)

# Add sample sizes
summary_stats$Sample_Size <- c(
  sum(!is.na(data$house_status_m)),
  sum(!is.na(data$house_status_m)),
  sum(!is.na(data$house_status_m)),
  sum(!is.na(data$house_burden_m)),
  sum(!is.na(data$house_moved_m)),
  sum(!is.na(data$house_payments_behind_m))
)

cat("\nKey Housing Metrics:\n")
print(knitr::kable(summary_stats,
      col.names = c("Metric", "Percentage (%)", "Sample Size (n)"),
      format = "simple"))

# Additional analysis of moving reasons for those who moved
if(sum(data$house_moved_m == "Yes", na.rm = TRUE) > 0) {
  total_movers <- sum(data$house_moved_m == "Yes", na.rm = TRUE)
  cat(sprintf("\nReasons for Moving (among those who moved) (n=%d):\n", total_movers))
  move_reasons <- data %>%
    filter(house_moved_m == "Yes") %>%
    filter(!is.na(house_move_reason_m)) %>%
    count(house_move_reason_m) %>%
    mutate(
      percentage = round(n/sum(n) * 100, 0)
    )
  print(knitr::kable(move_reasons,
        col.names = c("Reason", "Count", "Percentage (%)"),
        format = "simple"))
}
```
## Childcare {.tabset}
Analysis of childcare responsibilities, arrangements, and challenges.

```{r,}
# 1. Childcare Responsibilities
total_childcare_resp <- sum(!is.na(data$cc_has_children_m))
childcare_resp <- data %>%
  filter(!is.na(cc_has_children_m)) %>%
  count(cc_has_children_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nChildcare Responsibilities (n=%d):\n", total_childcare_resp))
print(knitr::kable(childcare_resp,
      col.names = c("Has Childcare Responsibilities", "Count", "Percentage (%)"),
      format = "simple"))

# 2. Age Range of Children - Modified for multiple selections
total_age_range_resp <- sum(!is.na(data$cc_age_range_m))

# Split and count age ranges
split_age_ranges <- function(x) {
  if(is.na(x)) return(NULL)
  ranges <- unlist(strsplit(x, ",\\s*"))
  return(trimws(ranges))
}

age_range_responses <- data$cc_age_range_m[!is.na(data$cc_age_range_m)]
all_age_ranges <- unlist(lapply(age_range_responses, split_age_ranges))

age_range <- data.frame(
  Age_Range = names(table(all_age_ranges)),
  Count = as.numeric(table(all_age_ranges))
) %>%
  mutate(
    percentage = round(Count/total_age_range_resp * 100, 0)  # Percentage based on total respondents
  ) %>%
  arrange(desc(Count))

cat(sprintf("\nAge Range of Children (n=%d respondents, multiple selections allowed):\n", 
            total_age_range_resp))
print(knitr::kable(age_range,
      col.names = c("Age Range", "Count", "Percentage of Respondents (%)"),
      format = "simple"))

# 3. Childcare Situation Changes
total_situation_changes <- sum(!is.na(data$cc_arrangement_m))
situation_changes <- data %>%
  filter(!is.na(cc_arrangement_m)) %>%
  count(cc_arrangement_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nChildcare Situation Changes (n=%d):\n", total_situation_changes))
print(knitr::kable(situation_changes,
      col.names = c("Change in Situation", "Count", "Percentage (%)"),
      format = "simple"))

# 4. Enrollment Status for Under 5
total_under5 <- sum(!is.na(data$cc_under5_care_m))
under5_enrollment <- data %>%
  filter(!is.na(cc_under5_care_m)) %>%
  count(cc_under5_care_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nEnrollment Status for Children Under 5 (n=%d):\n", total_under5))
print(knitr::kable(under5_enrollment,
      col.names = c("Enrollment Status", "Count", "Percentage (%)"),
      format = "simple"))

# For those who reported better childcare situation
if(sum(!is.na(data$childcare_better_m)) > 0) {
  total_better_resp <- sum(!is.na(data$childcare_better_m))
  
  # Function to split improvements
  split_improvements <- function(x) {
    if(is.na(x)) return(NULL)
    improvements <- unlist(strsplit(x, ",\\s*"))
    return(trimws(improvements))
  }
  
  # Process each response and count individual improvements
  better_responses <- data$childcare_better_m[!is.na(data$childcare_better_m)]
  all_improvements <- unlist(lapply(better_responses, split_improvements))
  
  better_situation <- data.frame(
    Improvement = names(table(all_improvements)),
    Count = as.numeric(table(all_improvements))
  ) %>%
    mutate(
      percentage = round(Count/total_better_resp * 100, 0)  # Percentage based on total respondents
    ) %>%
    arrange(desc(Count))
    
  cat(sprintf("\nImprovements in Childcare Situation (n=%d respondents, multiple selections allowed):\n", 
              total_better_resp))
  print(knitr::kable(better_situation,
        col.names = c("Type of Improvement", "Count", "Percentage of Respondents (%)"),
        format = "simple"))
}

# For those who reported worse childcare situation
if(sum(!is.na(data$childcare_worse_m)) > 0) {
  total_worse_resp <- sum(!is.na(data$childcare_worse_m))
  
  # Function to split responses while preserving commas within the same challenge
  split_challenges <- function(x) {
    if(is.na(x)) return(NULL)
    challenges <- unlist(strsplit(x, ",(?=[^,]+$)", perl=TRUE))
    return(trimws(challenges))
  }
  
  # Process each response and count individual challenges
  worse_responses <- data$childcare_worse_m[!is.na(data$childcare_worse_m)]
  all_challenges <- unlist(lapply(worse_responses, split_challenges))
  
  worse_situation <- data.frame(
    Challenge = names(table(all_challenges)),
    Count = as.numeric(table(all_challenges))
  ) %>%
    mutate(
      percentage = round(Count/total_worse_resp * 100, 0)  # Percentage based on total respondents
    ) %>%
    arrange(desc(Count))
    
  cat(sprintf("\nChallenges in Childcare Situation (n=%d respondents, multiple selections allowed):\n", 
              total_worse_resp))
  print(knitr::kable(worse_situation,
        col.names = c("Type of Challenge", "Count", "Percentage of Respondents (%)"),
        format = "simple"))
}

# Key Summary Statistics
summary_stats <- data.frame(
  Metric = c(
    "Members with Childcare Responsibilities",
    "Members with Children Under 5",
    "Members Reporting Better Childcare Situation",
    "Members Reporting Worse Childcare Situation"
  ),
  Percentage = c(
    round(mean(data$cc_has_children_m == "Yes", na.rm = TRUE) * 100, 0),
    round(mean(grepl("under 5", data$cc_age_range_m, ignore.case = TRUE), na.rm = TRUE) * 100, 0),
    round(mean(data$cc_arrangement_m == "My childcare situation is BETTER", na.rm = TRUE) * 100, 0),
    round(mean(data$cc_arrangement_m == "My childcare situation is WORSE", na.rm = TRUE) * 100, 0)
  )
)

# Add sample sizes to summary stats
summary_stats$Sample_Size <- c(
  sum(!is.na(data$cc_has_children_m)),
  sum(!is.na(data$cc_age_range_m)),
  sum(!is.na(data$cc_arrangement_m)),
  sum(!is.na(data$cc_arrangement_m))
)

cat("\nKey Childcare Metrics:\n")
print(knitr::kable(summary_stats,
      col.names = c("Metric", "Percentage (%)", "Sample Size (n)"),
      format = "simple"))
```
## Goals {.tabset}
Examination of members' goals, priorities, and progress.

```{r goals}
# 1. Overall Goals Analysis
total_goals_resp <- sum(!is.na(data$goals_m))
goals_overall <- data %>%
  filter(!is.na(goals_m)) %>%
  count(goals_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nPresence of Goals/Aspirations (n=%d):\n", total_goals_resp))
print(knitr::kable(goals_overall,
      col.names = c("Has Goals", "Count", "Percentage (%)"),
      format = "simple"))

# 2. Goal Priorities - Single clean output
total_priorities_resp <- sum(!is.na(data$goal_priority_m))

# Define the exact predefined categories
predefined_categories <- c(
    "Managing and improving my housing situation (e.g., covering costs, moving, making repairs)",
    "Changing my employment status (e.g., switching jobs, increasing/decreasing working hours, or starting a business)",
    "Continuing or starting education (e.g., a degree, certificate, or training)",
    "Paying down debt and/or catching up on bills",
    "Saving for the future",
    "Improving my transportation situation (e.g., car repairs, buying a car, etc.)",
    "Other (please specify)"
)

# Process the priorities
priorities_data <- data %>%
  filter(!is.na(goal_priority_m)) %>%
  mutate(goal_priority_m = as.character(goal_priority_m))

all_priorities <- vector("character")
for(response in priorities_data$goal_priority_m) {
    for(category in predefined_categories) {
        if(grepl(category, response, fixed=TRUE)) {
            all_priorities <- c(all_priorities, category)
        }
    }
}

# Create the summary table (single output)
priorities <- data.frame(
    Goal_Category = names(table(all_priorities)),
    Count = as.numeric(table(all_priorities))
) %>%
    mutate(
        percentage = round(Count/total_priorities_resp * 100, 0)
    ) %>%
    arrange(desc(Count))

# Single print of goal priorities
cat(sprintf("\nTop Goal Priorities (n=%d respondents, multiple selections allowed):\n", 
            total_priorities_resp))
print(knitr::kable(priorities,
      col.names = c("Goal Category", "Count", "Percentage of Respondents (%)"),
      format = "simple"))


# 3. Debt Goals Analysis - Modified for multiple selections
total_debt_resp <- sum(!is.na(data$goals_debt_m))

# Function to split and clean debt types
split_debt_types <- function(x) {
  if(is.na(x)) return(NULL)
  types <- unlist(strsplit(as.character(x), ",\\s*"))
  return(trimws(types))
}

# Process debt responses
debt_responses <- data$goals_debt_m[!is.na(data$goals_debt_m)]
all_debt_types <- unlist(lapply(debt_responses, split_debt_types))

debt_goals <- data.frame(
  Debt_Type = names(table(all_debt_types)),
  Count = as.numeric(table(all_debt_types))
) %>%
  mutate(
    percentage = round(Count/total_debt_resp * 100, 0)
  ) %>%
  arrange(desc(Count))

cat(sprintf("\nDebt Management Goals (n=%d respondents, multiple selections allowed):\n", total_debt_resp))
print(knitr::kable(debt_goals,
      col.names = c("Type of Debt", "Count", "Percentage of Respondents (%)"),
      format = "simple"))

# 4. Savings Goals - Modified for exact predefined categories
total_savings_resp <- sum(!is.na(data$goals_save_m))

# Define exact predefined savings categories
savings_categories <- c(
    "Retirement",
    "Buying a home",
    "Education (your own or someone else's)",
    "Leisure and Personal Well-being (e.g., hobbies, therapy, self-care)",
    "Emergency fund",
    "Health-related expenses",
    "Major purchases (like a car or appliances)",
    "Starting or expanding a business",
    "Supporting family needs (e.g., caregiving)",
    "Family vacations or activities",
    "Other (please specify)"
)

# Process savings responses with exact category matching
savings_responses <- data$goals_save_m[!is.na(data$goals_save_m)]
all_savings_goals <- vector("character")

for(response in savings_responses) {
    for(category in savings_categories) {
        if(grepl(category, response, fixed=TRUE)) {
            all_savings_goals <- c(all_savings_goals, category)
        }
    }
}

savings_goals <- data.frame(
    Savings_Goal = names(table(all_savings_goals)),
    Count = as.numeric(table(all_savings_goals))
) %>%
    mutate(
        percentage = round(Count/total_savings_resp * 100, 0)
    ) %>%
    arrange(desc(Count))

cat(sprintf("\nSavings Goals (n=%d respondents, multiple selections allowed):\n", total_savings_resp))
print(knitr::kable(savings_goals,
      col.names = c("Type of Savings Goal", "Count", "Percentage of Respondents (%)"),
      format = "simple"))

# 5. Key Summary Statistics
summary_stats <- data.frame(
  Metric = c(
    "Members with Specific Goals",
    "Members with Debt Management Goals",
    "Members with Savings Goals",
    "Members with Housing Goals",
    "Members with Employment Goals",
    "Members with Education Goals",
    "Members Finding Goals Easy to Achieve"
  ),
  Percentage = c(
    round(mean(data$goals_m == "Yes", na.rm = TRUE) * 100, 0),
    round(mean(!is.na(data$goals_debt_m) & data$goals_debt_m != "", na.rm = TRUE) * 100, 0),
    round(mean(!is.na(data$goals_save_m) & data$goals_save_m != "", na.rm = TRUE) * 100, 0),
    round(mean(!is.na(data$goal_housing_m) & data$goal_housing_m != "", na.rm = TRUE) * 100, 0),
    round(mean(!is.na(data$goal_employment_m) & data$goal_employment_m != "", na.rm = TRUE) * 100, 0),
    round(mean(!is.na(data$goal_education_m) & data$goal_education_m != "", na.rm = TRUE) * 100, 0),
    round(mean(grepl("easy", tolower(data$goals_debt_easy_hard_m)), na.rm = TRUE) * 100, 0)
  )
)

# Add sample sizes
summary_stats$Sample_Size <- c(
  sum(!is.na(data$goals_m)),
  sum(!is.na(data$goals_debt_m)),
  sum(!is.na(data$goals_save_m)),
  sum(!is.na(data$goal_housing_m)),
  sum(!is.na(data$goal_employment_m)),
  sum(!is.na(data$goal_education_m)),
  sum(!is.na(data$goals_debt_easy_hard_m))
)

cat("\nKey Goals and Financial Metrics:\n")
print(knitr::kable(summary_stats,
      col.names = c("Metric", "Percentage (%)", "Sample Size (n)"),
      format = "simple"))

table(data$goals_debt_easier_m)
table(data$goals_debt_harder_m)
table(data$goals_debt_easy_hard_m)

```

## Health {.tabset}
Analysis of physical health, mental health, and food security status.

```{r health}
# 1. General Health Status
total_health_resp <- sum(!is.na(data$wb_health_status_m))
health_status <- data %>%
  filter(!is.na(wb_health_status_m)) %>%
  count(wb_health_status_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nGeneral Health Status (n=%d):\n", total_health_resp))
print(knitr::kable(health_status,
      col.names = c("Health Status", "Count", "Percentage (%)"),
      format = "simple"))

# 2. Mental Health Analysis using both indicators
wellbeing_severity <- data %>%
  mutate(
    Anxiety = case_when(
      wb_anxious_m == "Nearly every day" | wb_worry_m == "Nearly every day" ~ "Severe",
      wb_anxious_m == "More than half the days" | wb_worry_m == "More than half the days" ~ "Moderate",
      wb_anxious_m == "Several days" | wb_worry_m == "Several days" ~ "Mild",
      TRUE ~ "None"
    ),
    Depression = case_when(
      wb_interest_m == "Nearly every day" | wb_down_m == "Nearly every day" ~ "Severe",
      wb_interest_m == "More than half the days" | wb_down_m == "More than half the days" ~ "Moderate",
      wb_interest_m == "Several days" | wb_down_m == "Several days" ~ "Mild",
      TRUE ~ "None"
    )
  ) %>%
  summarise(
    Anxiety_Severe = sum(Anxiety == "Severe", na.rm = TRUE),
    Anxiety_Moderate = sum(Anxiety == "Moderate", na.rm = TRUE),
    Anxiety_Mild = sum(Anxiety == "Mild", na.rm = TRUE),
    Anxiety_None = sum(Anxiety == "None", na.rm = TRUE),
    Depression_Severe = sum(Depression == "Severe", na.rm = TRUE),
    Depression_Moderate = sum(Depression == "Moderate", na.rm = TRUE),
    Depression_Mild = sum(Depression == "Mild", na.rm = TRUE),
    Depression_None = sum(Depression == "None", na.rm = TRUE)
  )

# Convert to long format for visualization
wellbeing_long <- wellbeing_severity %>%
  pivot_longer(cols = everything(), names_to = c("Condition", "Severity"), names_sep = "_") %>%
  mutate(
    Severity = factor(Severity, 
                     levels = c("None", "Mild", "Moderate", "Severe")),
    Condition = recode(Condition, 
                      "Anxiety" = "Anxiety",
                      "Depression" = "Depression")
  ) %>%
  group_by(Condition) %>%
  mutate(
    Percentage = round(value / sum(value) * 100, 0)
  ) %>%
  ungroup()

# Display detailed breakdown
cat("\nDetailed Mental Health Breakdown:\n")
print(knitr::kable(wellbeing_long %>% 
                   arrange(Condition, desc(Severity)),
      col.names = c("Condition", "Severity", "Count", "Percentage (%)"),
      format = "simple"))

# Create visualization
well_plot <- ggplot(wellbeing_long, aes(x = Severity, y = Percentage, fill = Severity)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  facet_wrap(~ Condition, scales = "free_y", ncol = 1) +
  scale_fill_manual(values = c("None" = "#dfd8e0", 
                              "Mild" = "#9f8da3", 
                              "Moderate" = "#664a6a", 
                              "Severe" = "#46214a")) +
  labs(x = "Severity", 
       y = "Percent of Members",
       title = sprintf("Mental Health Severity (n=%d)", 
                      sum(!is.na(data$wb_anxious_m)))) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1))

print(well_plot)

# 3. Food Insufficiency Status
total_food_resp <- sum(!is.na(data$wb_food_security_m))
food_insufficiency <- data %>%
  filter(!is.na(wb_food_security_m)) %>%
  count(wb_food_security_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)
  )

cat(sprintf("\nFood Insufficiency Status (n=%d):\n", total_food_resp))
print(knitr::kable(food_insufficiency,
      col.names = c("Food Insufficiency Status", "Count", "Percentage (%)"),
      format = "simple"))

# 4. Key Summary Statistics
summary_stats <- data.frame(
  Metric = c(
    "Good or Better Health",
    "Experiencing Moderate/Severe Anxiety",
    "Experiencing Moderate/Severe Depression",
    "Food Insufficiency",
    "Severe Food Insufficiency"
  ),
  Percentage = c(
    round(mean(data$wb_health_status_m %in% 
               c("Good", "Very good", "Excellent"), na.rm = TRUE) * 100, 0),
    round(mean(wellbeing_long$Percentage[wellbeing_long$Condition == "Anxiety" & 
                                       wellbeing_long$Severity %in% 
                                       c("Moderate", "Severe")]), 0),
    round(mean(wellbeing_long$Percentage[wellbeing_long$Condition == "Depression" & 
                                       wellbeing_long$Severity %in% 
                                       c("Moderate", "Severe")]), 0),
    round(mean(grepl("not enough", data$wb_food_security_m, 
                     ignore.case = TRUE), na.rm = TRUE) * 100, 0),
    round(mean(data$wb_food_security_m == 
               "Often not enough to eat", na.rm = TRUE) * 100, 0)
  )
)

# Add sample sizes
summary_stats$Sample_Size <- c(
  sum(!is.na(data$wb_health_status_m)),
  sum(!is.na(data$wb_anxious_m)),
  sum(!is.na(data$wb_down_m)),
  sum(!is.na(data$wb_food_security_m)),
  sum(!is.na(data$wb_food_security_m))
)

cat("\nKey Wellbeing Metrics:\n")
print(knitr::kable(summary_stats,
      col.names = c("Metric", "Percentage (%)", "Sample Size (n)"),
      format = "simple"))
```

## Health {.tabset}
Analysis of physical health, mental health, and food security status.

```{r health}
# 1. General Health Status (Health Perception)
total_health_resp <- sum(!is.na(data$wb_health_status_e))
health_status <- data %>%
  filter(!is.na(wb_health_status_e)) %>%
  count(wb_health_status_e) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0),
    # Create ordered factor for better sorting
    wb_health_status_e = factor(wb_health_status_e, 
                               levels = c("Excellent", "Very good", "Good", "Fair", "Poor"))
  ) %>%
  arrange(wb_health_status_e)

cat(sprintf("\nGeneral Health Status (n=%d):\n", total_health_resp))
print(knitr::kable(health_status,
      col.names = c("Health Status", "Count", "Percentage (%)"),
      format = "simple"))

# 2. Mental Health Analysis using both indicators
wellbeing_severity <- data %>%
  mutate(
    Anxiety = case_when(
      wb_anxious_e == "Nearly every day" | wb_worry_e == "Nearly every day" ~ "Severe",
      wb_anxious_e == "More than half the days" | wb_worry_e == "More than half the days" ~ "Moderate",
      wb_anxious_e == "Several days" | wb_worry_e == "Several days" ~ "Mild",
      TRUE ~ "None"
    ),
    Depression = case_when(
      wb_interest_e == "Nearly every day" | wb_down_e == "Nearly every day" ~ "Severe",
      wb_interest_e == "More than half the days" | wb_down_e == "More than half the days" ~ "Moderate",
      wb_interest_e == "Several days" | wb_down_e == "Several days" ~ "Mild",
      TRUE ~ "None"
    )
  ) %>%
  summarise(
    Anxiety_Severe = sum(Anxiety == "Severe", na.rm = TRUE),
    Anxiety_Moderate = sum(Anxiety == "Moderate", na.rm = TRUE),
    Anxiety_Mild = sum(Anxiety == "Mild", na.rm = TRUE),
    Anxiety_None = sum(Anxiety == "None", na.rm = TRUE),
    Depression_Severe = sum(Depression == "Severe", na.rm = TRUE),
    Depression_Moderate = sum(Depression == "Moderate", na.rm = TRUE),
    Depression_Mild = sum(Depression == "Mild", na.rm = TRUE),
    Depression_None = sum(Depression == "None", na.rm = TRUE)
  )

# Convert to long format for visualization
wellbeing_long <- wellbeing_severity %>%
  pivot_longer(cols = everything(), names_to = c("Condition", "Severity"), names_sep = "_") %>%
  mutate(
    Severity = factor(Severity, 
                     levels = c("None", "Mild", "Moderate", "Severe")),
    Condition = recode(Condition, 
                      "Anxiety" = "Anxiety",
                      "Depression" = "Depression")
  ) %>%
  group_by(Condition) %>%
  mutate(
    Percentage = round(value / sum(value) * 100, 0)
  ) %>%
  ungroup()

# Display detailed breakdown
cat("\nDetailed Mental Health Breakdown:\n")
print(knitr::kable(wellbeing_long %>% 
                   arrange(Condition, desc(Severity)),
      col.names = c("Condition", "Severity", "Count", "Percentage (%)"),
      format = "simple"))

# Create mental health visualization with brand colors
well_plot <- ggplot(wellbeing_long, aes(x = Severity, y = Percentage, fill = Severity)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  facet_wrap(~ Condition, scales = "free_y", ncol = 1) +
  scale_fill_manual(values = c("None" = "#FF8C00",      # Orange (positive)
                              "Mild" = "#5E2D91",       # Purple variant 
                              "Moderate" = "#D70073",   # Pink (moderate concern)
                              "Severe" = "#46214A")) +  # Main brand (severe concern)
  labs(x = "Severity", 
       y = "Percent of Members",
       title = sprintf("Mental Health Severity (n=%d)", 
                      sum(!is.na(data$wb_anxious_e)))) +
  theme_minimal() +
  theme(
    text = element_text(size = 12),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 14),
    plot.title = element_text(size = 16, hjust = 0.5),
    legend.position = "none",
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1))

print(well_plot)

# 3. Individual Mental Health Components Analysis
# Anxiety symptoms
total_anxiety_resp <- sum(!is.na(data$wb_anxious_e))
if(total_anxiety_resp > 0) {
  anxiety_breakdown <- data %>%
    filter(!is.na(wb_anxious_e)) %>%
    count(wb_anxious_e) %>%
    mutate(
      percentage = round(n/sum(n) * 100, 0)
    ) %>%
    arrange(desc(percentage))
  
  cat(sprintf("\nAnxiety Symptoms (Feeling nervous, anxious or on edge) (n=%d):\n", total_anxiety_resp))
  print(knitr::kable(anxiety_breakdown,
        col.names = c("Frequency", "Count", "Percentage (%)"),
        format = "simple"))
}

# Worry symptoms
total_worry_resp <- sum(!is.na(data$wb_worry_e))
if(total_worry_resp > 0) {
  worry_breakdown <- data %>%
    filter(!is.na(wb_worry_e)) %>%
    count(wb_worry_e) %>%
    mutate(
      percentage = round(n/sum(n) * 100, 0)
    ) %>%
    arrange(desc(percentage))
  
  cat(sprintf("\nWorry Symptoms (Not being able to stop or control worrying) (n=%d):\n", total_worry_resp))
  print(knitr::kable(worry_breakdown,
        col.names = c("Frequency", "Count", "Percentage (%)"),
        format = "simple"))
}

# Depression symptoms - lack of interest
total_interest_resp <- sum(!is.na(data$wb_interest_e))
if(total_interest_resp > 0) {
  interest_breakdown <- data %>%
    filter(!is.na(wb_interest_e)) %>%
    count(wb_interest_e) %>%
    mutate(
      percentage = round(n/sum(n) * 100, 0)
    ) %>%
    arrange(desc(percentage))
  
  cat(sprintf("\nLack of Interest (Little interest or pleasure in doing things) (n=%d):\n", total_interest_resp))
  print(knitr::kable(interest_breakdown,
        col.names = c("Frequency", "Count", "Percentage (%)"),
        format = "simple"))
}

# Depression symptoms - feeling down
total_down_resp <- sum(!is.na(data$wb_down_e))
if(total_down_resp > 0) {
  down_breakdown <- data %>%
    filter(!is.na(wb_down_e)) %>%
    count(wb_down_e) %>%
    mutate(
      percentage = round(n/sum(n) * 100, 0)
    ) %>%
    arrange(desc(percentage))
  
  cat(sprintf("\nDepressive Feelings (Feeling down, depressed, or hopeless) (n=%d):\n", total_down_resp))
  print(knitr::kable(down_breakdown,
        col.names = c("Frequency", "Count", "Percentage (%)"),
        format = "simple"))
}

# 4. Food Security Status
total_food_resp <- sum(!is.na(data$wb_food_security_e))
if(total_food_resp > 0) {
  food_security <- data %>%
    filter(!is.na(wb_food_security_e)) %>%
    count(wb_food_security_e) %>%
    mutate(
      percentage = round(n/sum(n) * 100, 0)
    ) %>%
    arrange(desc(percentage))
  
  cat(sprintf("\nFood Security Status (n=%d):\n", total_food_resp))
  print(knitr::kable(food_security,
        col.names = c("Food Security Status", "Count", "Percentage (%)"),
        format = "simple"))
}

# 5. Key Summary Statistics
summary_stats <- data.frame(
  Metric = c(
    "Good or Better Health",
    "Fair or Poor Health",
    "Experiencing Moderate/Severe Anxiety",
    "Experiencing Moderate/Severe Depression",
    "Food Insecurity (Any Level)",
    "Severe Food Insecurity",
    "Excellent Mental Health (No Symptoms)"
  ),
  Percentage = c(
    round(mean(data$wb_health_status_e %in% 
               c("Good", "Very good", "Excellent"), na.rm = TRUE) * 100, 0),
    round(mean(data$wb_health_status_e %in% 
               c("Fair", "Poor"), na.rm = TRUE) * 100, 0),
    ifelse(nrow(wellbeing_long) > 0,
           round(sum(wellbeing_long$value[wellbeing_long$Condition == "Anxiety" & 
                                        wellbeing_long$Severity %in% 
                                        c("Moderate", "Severe")]) / 
                 sum(wellbeing_long$value[wellbeing_long$Condition == "Anxiety"]) * 100, 0), 0),
    ifelse(nrow(wellbeing_long) > 0,
           round(sum(wellbeing_long$value[wellbeing_long$Condition == "Depression" & 
                                        wellbeing_long$Severity %in% 
                                        c("Moderate", "Severe")]) / 
                 sum(wellbeing_long$value[wellbeing_long$Condition == "Depression"]) * 100, 0), 0),
    round(mean(grepl("not enough|Sometimes not enough", data$wb_food_security_e, 
                     ignore.case = TRUE), na.rm = TRUE) * 100, 0),
    round(mean(grepl("Often not enough", data$wb_food_security_e, 
                     ignore.case = TRUE), na.rm = TRUE) * 100, 0),
    ifelse(nrow(wellbeing_long) > 0,
           round(mean(wellbeing_long$value[wellbeing_long$Severity == "None"]) / 
                 sum(wellbeing_long$value) * 100, 0), 0)
  )
)

# Add sample sizes
summary_stats$Sample_Size <- c(
  sum(!is.na(data$wb_health_status_e)),
  sum(!is.na(data$wb_health_status_e)),
  sum(!is.na(data$wb_anxious_e)),
  sum(!is.na(data$wb_down_e)),
  sum(!is.na(data$wb_food_security_e)),
  sum(!is.na(data$wb_food_security_e)),
  sum(!is.na(data$wb_anxious_e))
)

cat("\nKey Health and Wellbeing Metrics:\n")
print(knitr::kable(summary_stats,
      col.names = c("Metric", "Percentage (%)", "Sample Size (n)"),
      format = "simple"))

# 6. Health Status Categories Summary
if(total_health_resp > 0) {
  positive_health <- sum(data$wb_health_status_e %in% c("Excellent", "Very good", "Good"), na.rm = TRUE)
  concerning_health <- sum(data$wb_health_status_e %in% c("Fair", "Poor"), na.rm = TRUE)
  
  cat(sprintf("\nOverall Health Status Summary:\n"))
  cat(sprintf("- Positive Health (Excellent/Very good/Good): %d (%d%%)\n", 
              positive_health, round(positive_health/total_health_resp * 100, 0)))
  cat(sprintf("- Concerning Health (Fair/Poor): %d (%d%%)\n", 
              concerning_health, round(concerning_health/total_health_resp * 100, 0)))
}
```
## Community and Social Support {.tabset}
Overview of community participation and support networks.
```{r community-social}
# 1. Community Participation Analysis
clean_activities <- function(x) {
  activities <- c(
    "Religious Activities",
    "Community Betterment",
    "Health and Wellbeing Activities",
    "Educational Involvement",
    "Civic Activities",
    "Creative and Cultural Activities",
    "Neighborhood Groups",
    "Local Business Support",
    "Recreational Groups",
    "Other",
    "No Participation"
  )
  
  all_responses <- unlist(strsplit(as.character(x), ","))
  
  categorized <- sapply(all_responses, function(resp) {
    if(grepl("Religious", resp)) return("Religious Activities")
    if(grepl("Community Betterment", resp)) return("Community Betterment")
    if(grepl("Health", resp)) return("Health and Wellbeing Activities")
    if(grepl("Educational", resp)) return("Educational Involvement")
    if(grepl("Civic", resp)) return("Civic Activities")
    if(grepl("Creative", resp)) return("Creative and Cultural Activities")
    if(grepl("Neighborhood", resp)) return("Neighborhood Groups")
    if(grepl("Local Business", resp)) return("Local Business Support")
    if(grepl("Recreational", resp)) return("Recreational Groups")
    if(grepl("Other", resp)) return("Other")
    if(grepl("not participated", resp)) return("No Participation")
    return(NA)
  })
  
  return(categorized)
}

# Community Activity Analysis
total_respondents <- sum(!is.na(data$comm_part_m))
activity_counts <- table(unlist(lapply(data$comm_part_m, clean_activities)))
activity_df <- data.frame(
  Activity = names(activity_counts),
  Count = as.numeric(activity_counts)
) %>%
  mutate(
    Percentage = round(Count/total_respondents * 100, 0)  # Round to whole numbers
  ) %>%
  arrange(desc(Count))

cat(sprintf("\nCommunity Activity Participation (n=%d):\n", total_respondents))
print(knitr::kable(activity_df,
      col.names = c("Type of Activity", "Count", "Percentage (%)"),
      format = "simple"))

# 2. Support Type Analysis
analyze_support <- function(data, column, support_type) {
  total_resp <- sum(!is.na(data[[column]]))
  freq_table <- table(data[[column]], useNA = "ifany")
  support_df <- data.frame(
    Frequency = names(freq_table),
    Count = as.numeric(freq_table)
  ) %>%
    mutate(
      Percentage = round(Count/sum(Count) * 100, 0)  # Round to whole numbers
    )
  
  cat(sprintf("\n%s (n=%d):\n", support_type, total_resp))
  print(knitr::kable(support_df,
        col.names = c(paste(support_type, "Frequency"), "Count", "Percentage (%)"),
        format = "simple"))
}

# Define support types
support_types <- list(
  "Emotional Support" = "comm_give_emo_supp_m",
  "Financial Support" = "comm_give_money_m",
  "Caregiving Support" = "comm_give_caregiving_m",
  "Knowledge Sharing" = "comm_give_knowledge_m",
  "Transportation Support" = "comm_give_transport_m",
  "Errands Support" = "comm_give_errands_m",
  "Food Support" = "comm_give_food_m"
)

# Run analysis for each support type
for(support_name in names(support_types)) {
  analyze_support(data, support_types[[support_name]], support_name)
}

# 3. Calculate regular support summary
calculate_regular_support <- function(data, column) {
  sum(data[[column]] %in% c("Often", "Sometimes"), na.rm = TRUE) / 
    sum(!is.na(data[[column]])) * 100
}

regular_support <- sapply(support_types, function(col) {
  calculate_regular_support(data, col)
})

support_summary <- data.frame(
  Support_Type = names(support_types),
  Regular_Support = round(regular_support, 0),  # Round to whole numbers
  Sample_Size = sapply(support_types, function(col) sum(!is.na(data[[col]])))
) %>%
  arrange(desc(Regular_Support))

cat("\nRegular Support Summary (Often + Sometimes):\n")
print(knitr::kable(support_summary,
      col.names = c("Type of Support", "Percentage Providing Regular Support (%)", "Sample Size (n)"),
      format = "simple"))

# Community Support relationships
total_relationships <- sum(!is.na(data$comm_relationships_m))
relationships_df <- data %>%
  filter(!is.na(comm_relationships_m)) %>%
  count(comm_relationships_m) %>%
  mutate(
    percentage = round(n/sum(n) * 100, 0)  # Round to whole numbers
  ) %>%
  arrange(desc(n))

cat(sprintf("\nCommunity Relationships Strength (n=%d):\n", total_relationships))
print(knitr::kable(relationships_df,
      col.names = c("Relationship Level", "Count", "Percentage (%)"),
      format = "simple"))
```
## Survey Feedback {.tabset}
Analysis of member feedback and suggestions.
```{r survey-feedback}
library(wordcloud)
library(tm)

# Create Spanish to English translation dictionary
spanish_to_english <- list(
  # Basic terms
  "gracias" = "thank you",
  "ayuda" = "help",
  "familia" = "family",
  "dinero" = "money",
  "apoyo" = "support",
  "programa" = "program",
  "necesidad" = "need",
  "trabajo" = "work",
  "pago" = "payment",
  "casa" = "house",
  "hogar" = "home",
  "comida" = "food",
  "niños" = "children",
  "mejor" = "better",
  "bueno" = "good",
  "excelente" = "excellent",
  "mucho" = "much",
  "mas" = "more",
  "tiempo" = "time",
  "vida" = "life",
  "continuar" = "continue",
  "ayudar" = "help",
  "personas" = "people",
  "comunidad" = "community",
  "beneficio" = "benefit",
  "muchas" = "many",
  "seguir" = "continue",
  "pregunta" = "question",
  "respuesta" = "response",
  "experiencia" = "experience"
  # Add more translations as needed
)

# Function to count Spanish responses
count_spanish_responses <- function(text_vector) {
  # Count responses containing common Spanish words
  spanish_indicators <- c("gracias", "ayuda", "familia", "este", "por", "para", "con", "los", "las", "el", "la")
  spanish_count <- sum(sapply(text_vector, function(x) 
    any(sapply(spanish_indicators, function(word) 
      grepl(word, tolower(x))))))
  return(spanish_count)
}

# Function to translate Spanish text to English
translate_text <- function(text) {
  text_lower <- tolower(text)
  for(span_word in names(spanish_to_english)) {
    text_lower <- gsub(span_word, spanish_to_english[[span_word]], text_lower)
  }
  return(text_lower)
}

# Prepare and analyze feedback
feedback_data <- data %>%
  filter(!is.na(survey_feedback_m))

# Count Spanish responses before translation
spanish_response_count <- count_spanish_responses(feedback_data$survey_feedback_m)

# Translate and prepare text for word cloud
translated_feedback <- sapply(feedback_data$survey_feedback_m, translate_text)

# Create corpus with translated text
feedback_corpus <- Corpus(VectorSource(translated_feedback))
feedback_corpus <- tm_map(feedback_corpus, removePunctuation)
feedback_corpus <- tm_map(feedback_corpus, removeNumbers)
feedback_corpus <- tm_map(feedback_corpus, removeWords, stopwords("english"))
feedback_corpus <- tm_map(feedback_corpus, stripWhitespace)

# Create document term matrix
dtm <- TermDocumentMatrix(feedback_corpus)
matrix <- as.matrix(dtm)
words <- sort(rowSums(matrix), decreasing=TRUE)
df <- data.frame(word = names(words), freq=words)

# Generate word cloud
print("Survey Feedback Word Cloud:")
wordcloud(words = df$word, freq = df$freq, min.freq = 2,
          max.words=100, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

# Print top themes
print("\nMost Common Themes in Feedback:")
top_themes <- head(df, 20)
print(knitr::kable(top_themes,
      col.names = c("Theme", "Frequency"),
      format = "simple"))

# Print translation note
cat(sprintf("\nNote: %d feedback responses were translated from Spanish to English for standardization in analysis.\n", 
            spanish_response_count))

# Optional: Show some example translations
print("\nExample Translations (Original → Translated):")
# Show a few examples where translation occurred
translated_examples <- data.frame(
  Original = feedback_data$survey_feedback_m[1:3],
  Translated = translated_feedback[1:3]
)
print(knitr::kable(translated_examples,
      col.names = c("Original Response", "Translated Response"),
      format = "simple"))


```



